### 2014-08-03 tex/GAME_HARD_01_BIOGRAPHY.tex
Added email-descriptor to the e-mail address under section Biography.

% GAME_HARD_01_BIOGRAPHY.tex

\section{AUTHOR BIOGRAPHY}
\label{sec:authorbiography}
\textbf{ERIC NILSSON} conducted his master thesis in Computer Science on graphics acceleration in virtual platforms ('Paravirtualizing OpenGL ES in Simics') in collaboration with Intel Corporation to obtain his M.Sc.E. from Blekinge Institute of Technology, where he majors in Software~Engineering.
In addition to his interest in the sciences, he has a keen interest in horror games; an engagement he pursues to the best of his ability.
Currently, Eric is involved with system simulation at Intel where he carries out his duties as a Software Engineer.\\

E-mail: \href{mailto:erne09@student.bth.se}{erne09@student.bth.se}