### 2014-08-24 presentation/GAME_HARD_01_PRESENTATION_CONCLUSION.tex
Added presentation outline.

% GAME_HARD_01_PRESENTATION_CONCLUSION.tex
\section{Conclusion}

% SUBSECTION
\subsection{Subsection}
\begin{frame}
\frametitle{SUBSECTION}

\end{frame}