### 2014-09-07 presentation/GAME_HARD_01_PRESENTATION_CONCLUSION.tex
Added draft results tables to section Conclusion.

% GAME_HARD_01_PRESENTATION_CONCLUSION.tex
\section{Conclusion}

% Arrowhead command. Sets a tikz node for rendering an entire arrow.
\newcommand\ah[2]{\tikz[remember picture,baseline]\node[inner sep=2pt,outer sep=0](#1){#2};}
% Arrow command. Draws an arrow between two tikz nodes.
\newcommand\arrow[2]{%
  \begin{tikzpicture}[remember picture, overlay, >=stealth, shift={(0,0)}]
    \draw[->] (#1) to (#2);
  \end{tikzpicture}%
}

% RESULTS
\subsection{Results}
\begin{frame}
\frametitle{RESULTS}

\begin{columns}[T]
  \begin{column}{\widthof{I}}
    \begin{sideways}
      Integer
    \end{sideways}
  \end{column}
  \begin{column}{\textwidth - \widthof{I}}
    \begin{center}
      \begin{tabular}{r|r|r|r|}
        \cline{2-3}
        & \multicolumn{1}{|c|}{\textbf{BASIC}} & \multicolumn{1}{|c|}{\textbf{TILED}} \\ \hline
        \multicolumn{1}{|l|}{\textbf{HARD}}     & \ah{hh1}{$1.16$}      & \ah{hf1}{\phantom{0000}$0.24$}\ah{spc}{}      & \ah{spc}{$-79.3\%$} \\ \hline \hline
        \multicolumn{1}{|l|}{\textbf{SOFT}}     & \ah{vh2}{}\ah{hh2}{$11610.02$}        & \ah{hf2}{\phantom{0}$9866.40$}\ah{vh1}{}      & \ah{spc}{$-15.0\%$} \\ \hline
        \multicolumn{1}{|l|}{\textbf{WARP}}     & \ah{vf2}{}\phantom{000}\ah{hh3}{$15.31$}      & \ah{hf3}{\phantom{000}$18.97$}\ah{vf1}{}      & \ah{spc}{$+23.9\%$} \\ \hline
        & \multicolumn{1}{|r|}{\ah{spc}{$-99.9\%$}} & \multicolumn{1}{|r|}{\ah{spc}{$-99.8\%$}} \\ \cline{2-3}
      \end{tabular}
      \arrow{hh1}{hf1} % Draw horizontal arrow heads to fletchlings.
      \arrow{hh2}{hf2}
      \arrow{hh3}{hf3}
      \arrow{vh1}{vf1} % Draw vertical arrow heads to fletchlings.
      \arrow{vh2}{vf2}
    \end{center}
  \end{column}
\end{columns}

\begin{columns}[T]
  \begin{column}{\widthof{I}}
    \begin{sideways}
      Floating point
    \end{sideways}
  \end{column}
  \begin{column}{\textwidth - \widthof{I}}
    \begin{center}
      \begin{tabular}{r|r|r|r|}
        \cline{2-3}
        & \multicolumn{1}{|c|}{\textbf{BASIC}} & \multicolumn{1}{|c|}{\textbf{TILED}} \\ \hline
        \multicolumn{1}{|l|}{\textbf{HARD}}     & \ah{hh1}{$0.77$}      & \ah{hf1}{\phantom{0000}$0.22$}\ah{spc}{}      & \ah{spc}{$-71.4\%$} \\ \hline \hline
        \multicolumn{1}{|l|}{\textbf{SOFT}}     & \ah{vh1}{}\ah{hh2}{$10247.03$}        & \ah{hf2}{$10909.88$}\ah{vh2}{}        & \ah{spc}{$+6.5\%$} \\ \hline
        \multicolumn{1}{|l|}{\textbf{WARP}}     & \ah{vf1}{}\phantom{000}\ah{hh3}{$14.08$}      & \ah{hf3}{\phantom{000}$17.44$}\ah{vf2}{}      & \ah{spc}{$+23.9\%$} \\ \hline
        & \multicolumn{1}{|r|}{\ah{spc}{$-99.9\%$}} & \multicolumn{1}{|r|}{\ah{spc}{$-99.8\%$}} \\ \cline{2-3}
      \end{tabular}
      \arrow{hh1}{hf1} % Draw horizontal arrow heads to fletchlings.
      \arrow{hh2}{hf2}
      \arrow{hh3}{hf3}
      \arrow{vh1}{vf1} % Draw vertical arrow heads to fletchlings.
      \arrow{vh2}{vf2}
    \end{center}
  \end{column}
\end{columns}

\end{frame}

% Results

% Impact of Shared Memory

% Future Work

% Source code and data available
% May be downloaded if you want to replicate the test