### 2014-09-07 presentation/GAME_HARD_01_PRESENTATION_CONCLUSION.tex
Moved results and shared memory impacts to section Contribution.

% GAME_HARD_01_PRESENTATION_CONCLUSION.tex

% CONCLUSION
\section{Conclusion}
\begin{frame}
\frametitle{CONCLUSION}

Conclusions here

\end{frame}

% FUTURE WORK
\subsection{Future Work}
\begin{frame}
\frametitle{FUTURE WORK}

\begin{itemize}
\item Memory bandwidth
\end{itemize}

\end{frame}