### 2014-09-08 presentation/GAME_HARD_01_PRESENTATION_CONCLUSION.tex
Authored draft conclusion material under section CONCLUSION.

% GAME_HARD_01_PRESENTATION_CONCLUSION.tex

% CONCLUSION
\section{Conclusion}
\begin{frame}
\frametitle{CONCLUSION}

\begin{itemize}
\item Results indicate a major performance boost using WARP, when compared to that of the reference device driver
  \begin{itemize}
  \item Accounting for roughly three orders of magnitude
  \end{itemize}
\item However, important to keep in mind that the drivers may be useful for different purposes
  \begin{itemize}
  \item This calls into question what the flaws of using WARP, in place of the reference device driver, may be
  \end{itemize}
\end{itemize}

\end{frame}

% FUTURE WORK
\subsection{Future Work}
\begin{frame}
\frametitle{FUTURE WORK}

% Investigate potential memory bandwidth bottleneck
% Establish what the flaws of using WARP are
% Potential use-case: the acceleration of graphics in some virtual platforms
% Double floating point precision
% Complex kernel experiment

\end{frame}