### 2014-09-08 presentation/GAME_HARD_01_PRESENTATION_CONCLUSION.tex
Authored draft future work material under section CONCLUSION.

% GAME_HARD_01_PRESENTATION_CONCLUSION.tex

% CONCLUSION
\section{Conclusion}
\begin{frame}
\frametitle{CONCLUSION}

\begin{itemize}
\item Results indicate a major performance boost using WARP, when compared to that of the reference device driver
  \begin{itemize}
  \item Accounting for roughly three orders of magnitude
  \end{itemize}
\item However, important to keep in mind that the drivers may be useful for different purposes
  \begin{itemize}
  \item This calls into question what the flaws of using WARP, in place of the reference device driver, may be
  \end{itemize}
\end{itemize}

\end{frame}

% FUTURE WORK
\subsection{Future Work}
\begin{frame}
\frametitle{FUTURE WORK}

\begin{enumerate}
\item Microsoft speculates memory bandwidth as a potential bottleneck for throughput-oriented workloads on latency-oriented systems
\item Investigate the performance of WARP in coagency with more complex workloads
\item For brevity, establish whether or not there may be precisional discrepency in double floating point precision
\end{enumerate}

A potential use-case worthy of investigating may be using WARP-like technologies for the purposes of accelerating graphics in virtual platforms.

\end{frame}